Spatial Relational Models for Neural Circuit Discovery

### 2013-10-21 docs/techreport.tex
updates to the indicated plots tech report progressing

\documentclass{article}
\usepackage[english,american]{babel}
\usepackage{todonotes}
\usepackage{subfiles}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{fullpage}

@@

\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{multicol}

@@

\subsubsection{Structural Inference}
Algorithm 8 from radford neal \autocite{Neal2000}

@@

\subsubsection{Hyperparameter Inferece}
Gridded gibbs sampling for hyperparameter inference

\subsubsection{Annealing}
For many examples of the data here we perform annealing.

@@

Dense serial electron microscopy of a $VOLUMNE$ in the mouse
retina by \autocite{Helmstaedter2013} yielded a listing of places where
neurons come into contact. There were $number$ cells originally, whcih
were refined through various classification schemes. We selected the
$950$ for which the location of the soma could be reconstruted from
the provided cell plots (soma locations were not provided by the
study's authors in machien-readble form).

@@

\begin{figure}
  \centering 
  \includegraphics[width=0.5\textwidth]{mouseretina/adjmat.byz.png}
  \caption{Mouse connectome for the 950 cells with known soma positions, ordered by the position of the soma along the $z$ axis. The size of each point is proportional to the total area of contact between the cells.}
  \label{fig:mouse_adj}
\end{figure}

\subsection{C. elegans}
Connectome across several animals traced out by sidney brenner back in the day.
Small number of cells
position is linear along body axis
cell divisions are really precise

only look at hermaphrdite, ignore phyranx.

Have directionality, chemical vs electrical, number

279 neurons (non-phyrnx)

\begin{figure}
  \centering 
  \includegraphics[width=0.5\textwidth]{celegans/data.pos.pdf}
  \caption{Connectome for the 279 non-pharynx cells in \textit{c. elegans}. Cells are ordered by their position along the body axis. Blue dots are chemical synapses, red dots are electrical synapses, and the area of each dot is equivalent to the number of synaptic contacts between the cells. Note that the chemical synapses (blue) are directional, thus the red dots form a symmetric matrix}
  \label{fig:celegans_adj}
\end{figure}

@@

\begin{figure}
  \centering 
  \includegraphics[width=0.5\textwidth]{drosophila/adj_comp.pdf}
  \caption{Connectome for the 357 cells in \textit{drosophila} medulla. Cells are ordered by the order they appear in the data file. Dot area is proportional to the number of synapses; data are directional.}
  \label{fig:drosophila_adj}
\end{figure}

@@

>>>>

\documentclass{article}
\usepackage[english,american]{babel}
\usepackage{graphicx}
\usepackage{todonotes}
\usepackage{subfiles}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{fullpage}

@@

\usepackage{amsmath}
\usepackage{subcaption}

\usepackage[firstpage]{draftwatermark}
\SetWatermarkScale{4}

@@

\subsubsection{Structural Inference}
Algorithm 8 from radford neal \autocite{Neal2000} for dirichlet process
inference.

@@

\subsubsection{Hyperparameter Inferece}
Gridded gibbs sampling for hyperparameter inference
We set the grid by hand (empirical bayes)

\subsubsection{Annealing}
For many examples of the data here we perform annealing on the global likelihood during the burn-in period.

@@

Dense serial electron microscopy of a $VOLUMNE$ in the mouse
retina by \autocite{Helmstaedter2013} yielded a listing of places where
neurons come into contact. There were $number$ cells originally, whcih
were refined through various classification schemes. We selected the
$950$ for which the location of the soma could be reconstruted from
the provided cell plots (soma locations were not provided by the
study's authors in machine-readble form).

@@

\begin{figure}
  \centering 
  \includegraphics[width=0.5\textwidth]{mouseretina/adjmat.byz.png}
  \caption{Mouse connectome for the 950 cells with known soma positions, ordered by the position of the soma along the $z$ axis. The size of each point is proportional to the total area of contact between the cells.}
  \label{fig:mouseretina:adj}
\end{figure}

\subsubsection{Results}

\begin{figure}
  \centering 
  \includegraphics[width=\textwidth]{mouseretina/retina.1.0.ld.0.0.data-fixed_20_100-anneal_slow_400.0.clusters.pdf}
  \caption{Identified clusters from the mouse retina connectome using a Logistic-Distance-Bernoulli likelihood. Each column of plots is a cluster, wider plots for clusters with more cells. \textbf{first row} of plots compares the identified cluster to the experimenter-provided latent cell types, of which there were 74. Each bar is the fraction (between 0 and 1) of the cells of that type that are in this cluster. Note that the experimentors ordered their latent cell types, so cell type 60 is closer to 61 than 65 (see main text). Colors reflect the macro cell types (red = retinal ganglion cells, blue are amacrine, green are bipolar, and black are other. \textbf{Second row} is the spatial distribution of soma (blue) and synapses (red) for this cluster along a direction approximately normal to the retina surface. \textbf{Third row} is the spatial distribution of the soma (blue dots) and synapses (red dots) in this cluster, across the retina surface.}
  \label{fig:mouseretina:clusters}
\end{figure}

\begin{figure}
  \centering 
  \begin{subfigure}[b]{0.46\textwidth}
    \includegraphics[width=\textwidth, page=1]{mouseretina/retina.1.0.ld.0.0.data-fixed_20_100-anneal_slow_400.0.latent.pdf}
    \caption{Clustered latent connectivity matrix.}
    \label{fig:gull}
  \end{subfigure}
  \begin{subfigure}[b]{0.46\textwidth}
    \includegraphics[width=\textwidth, page=2]{mouseretina/retina.1.0.ld.0.0.data-fixed_20_100-anneal_slow_400.0.latent.pdf}
    \caption{Latent cluster parameters. }
    \label{fig:gull}
  \end{subfigure}
  \caption{Latent structure for mouse retina connectome clusters above. The latent cluster parameters in (b.) show (blue) the empirical distribution of connectivity for this cluster as a function of distance, and (red) the fit logistic curve. The vertical black line is the value of $\mu$.}
  \label{fig:mouseretina:latent}
\end{figure}

\subsection{C. elegans}

The connectome of the nematode Caenorhabditis elegans was traced by
hand from serial electron micrographs from Sidney Brenner's laboratory
in 1986 \autocite{White1986}. \textit{C. elegans} exhibits extremely
stereotypic neural development, with each hermaphrodite animal having
exactly 302 neurons, and the synaptic connectivity being 75\%
identical \todo{where did this number come from} across animals. Here we
examine 279 nonpharyngeal neurons with 6393 chemical synapes and 890 electrical junctions originally cleaned up in \autocite{Chen2006}. The adjacency matrix is shown in figure~\ref{fig:celegans:adj}

\begin{figure}
  \centering 
  \includegraphics[width=0.5\textwidth]{celegans/data.pos.pdf}
  \caption{Connectome for the 279 nonpharyngeal cells in \textit{c. elegans}. Cells are ordered by their position along the body axis. Blue dots are chemical synapses, red dots are electrical synapses, and the area of each dot is equivalent to the number of synaptic contacts between the cells. Note that the chemical synapses (blue) are directional, thus the red dots form a symmetric matrix}
  \label{fig:celegans:adj}
\end{figure}

The structure of the \textit{c. elegans} nervous system differs considerably 
from the primary sensory connectomes we examined:

\begin{itemize}
\item Lots of ganglia with global projections
\item Many of Brenner's originally-identified cell types contain only two member cells, one on each side of the animal. 
\item We have both electrical and chemical synapses, where the chemical synapses are directional. 
\item Our position data only indicates the position along the body axis (on the interval $[0, 1]$. 
\end{itemize}

The neurons can be broadly classified into ``sensory'', ``motor'', and
``interneurons''. In addition, there are several broad classes of both
sensory and motor neurons containing up to twelve cells. Figure ~\ref{fig:celegans:classestruth} shows the locations the cells of the biggest classes. Note the anterior concentration of sensory and internurons in contrast to the broad spatial distribution of the motor neurons.

\begin{figure}
  \centering 
  \includegraphics[width=0.8\textwidth]{celegans/class.positions.pdf}
  \caption{Positions of \textit{C. elegans} neurons grouped by cell type along the anterior-posterior axis. The cell type is listed to the right. Blue cells are motor neurons, green are sensory, and grey are interneurons. }
  \label{fig:celegans:classestruth}
\end{figure}

\subsubsection{results}
They are crappy.

\todo{Should we try to include the neuromuscular and sensory junctions? The data exists from wormatlas.org}

@@

\begin{figure}
  \centering 
  \includegraphics[width=0.7\textwidth]{drosophila/adj_comp.pdf}
  \caption{Connectome for the 357 cells in \textit{drosophila} medulla. Cells are ordered by the order they appear in the data file. Dot area is proportional to the number of synapses; data are directional.}
  \label{fig:drosophila_adj}
\end{figure}

\subsubsection{results}

@@

\subsubsection{results}
